---
title: "Analysis of the used cars dataset"
author: ""
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Data Import and Cleaning
```{r cars, include=FALSE}

df = read.csv("cars.csv", header = T, stringsAsFactors = TRUE)

df = subset(df, select=c("manufacturer_name", "model_name", "transmission", "color", "odometer_value", "year_produced", "engine_fuel", "engine_has_gas", "engine_type", "engine_capacity", "body_type", "has_warranty",
"state", "drivetrain", "price_usd", "is_exchangeable", "number_of_photos",  "up_counter"))
names(df)[names(df) == 'price_usd'] <- 'price'

```




## Renaming features
```{r include=FALSE}
names(df)[names(df) == 'price_usd'] <- 'price'
```


Checking for the the factors and numerical features   
```{r include=FALSE}
str(df)
```
Looks perfectly fine

## Data Preprocessing
```{r include=FALSE}
colSums(is.na(df))
```

Engine Capacity has 10 null values, we dropped the rows with null values.
```{r include=FALSE}
#Droping the rows with null values 
df <- df[complete.cases(df), ]
```

## Summary of Dataset
```{r include=FALSE}
library(corrplot)
library(tidyverse)

df %>% select_if(is.numeric)->cars_numerical
```


•	The Dataset used for the project is “Belarus-Used-cars-catalog” taken from the public data source Kaggle (An online community of data scientists and machine learning practitioners).

  Link: https://www.kaggle.com/lepchenkov/usedcarscatalog?select=cars.csv

•	The Dataset contains information about the Belarus (western Europe) used cars market from the year 2019.

•	The total number of variables in the data set is 19.

•	The total number of observations in the data set is 38521.

•	This Dataset helps the team in exploring the used car market in Belarus and build a model to find the relationship between car prices with changing features that can effectively predict the price of a used car, given the certain parameters (both numerical and categorical).

•	From the Dataset we mainly focus on these features as mentioned below to perform Exploratory Data Analysis:

  •	Color
  •	Transmission
  •	Odometer value
  •	Year of Production
  •	Body type
  •	Number of Photos
  •	Duration of days


### Limitations of Dataset:
1.	The “Belarus-Used-cars-catalog” dataset is limited to only Belarus which in effect does not help Pattern 6 to make assumptions about used car markets in other countries.

2.	There is no ‘electric’ car category as the dataset is limited to gasoline and diesel.

3.	There could have been more features found in the dataset which Team Pattern 6 could have used for the Exploratory Data Analysis and get a more detailed analysis when comparing multiple features.



## Exploratory Data Analysis
```{r echo=TRUE}
# summary(cars_numerical)
library(fBasics)
basicStats(cars_numerical)
```

The table above gives us the basic statistic measures of numeric data. There are six numerical variables in our dataset. The one we care about most is the used car's price. It has mean=6640, standard deviation(sd)=6430. The odometer_value with mean=249000, sd=136000. The year_produced with mean=2000 and sd=8.06. The engine_capacity has mean=2.06 and sd=0.67. The absolute values of skewness for all the variables are all greater than 1, which indicates they are highly skewed. The kurtosis values are all greater than 0, indicating they are sharply peaked with heavy tails. More analysis between other variables shows below.

## Normality tests

This section checks the normality of numerical variables based on the Q-Q plot, histogram, and normality tests. The most common method for normality test is called *Shapiro-Wilk's method*, however, this test only works when the observation is less than 5000 in __R__,and our dataset is more extensive than this value, so we will use the *Kolmogorov-Smirnov (K-S) normality test* instead.

```{r, fig.height=3}
library(gridExtra)
plot1 = ggplot(cars_numerical, aes(sample = price)) + stat_qq(col="#00AFBB") + stat_qq_line() + labs(title = 'Q-Q plot of price') 
plot2 = ggplot(cars_numerical, aes(x = price)) + geom_histogram(fill = "#00AFBB", colour="white", bins=40) + labs(title = 'Histogram of price')

grid.arrange(plot1, plot2, ncol=2, nrow=1)
```

As we can find in the quantile-quantile plot and the histogram,`price` is not normally distributed, if we want use price as the dependent variable for a linear regression, it is necessary to transform it to a normal distribution after that.

```{r}
plot3 = ggplot(cars_numerical, aes(sample = odometer_value)) + stat_qq(col="#00AFBB") + stat_qq_line() + labs(title = 'Q-Q plot of odometer_value')
plot4 = ggplot(cars_numerical, aes(x = odometer_value)) + geom_histogram(fill = "#00AFBB", colour="white", bins=40) + labs(title = 'Histogram of odometer_value')

plot5 = ggplot(cars_numerical, aes(sample = year_produced)) + stat_qq(col="#00AFBB") + stat_qq_line() + labs(title = 'Q-Q plot of year_produced')
plot6 = ggplot(cars_numerical, aes(x = year_produced)) + geom_histogram(fill = "#00AFBB", colour="white", bins=40) + labs(title = 'Histogram of year_produced')

grid.arrange(plot3, plot4, plot5, plot6, ncol=2, nrow=2)
```

```{r}
plot7 = ggplot(cars_numerical, aes(sample = engine_capacity)) + stat_qq(col="#00AFBB") + stat_qq_line() + labs(title = 'Q-Q plot of engine_capacity')
plot8 = ggplot(cars_numerical, aes(x = engine_capacity)) + geom_histogram(fill = "#00AFBB", colour="white", bins=40) + labs(title = 'Histogram of engine_capacity')

plot9 = ggplot(cars_numerical, aes(sample = number_of_photos)) + stat_qq(col="#00AFBB") + stat_qq_line() + labs(title = 'Q-Q plot of number_of_photos')
plot10 = ggplot(cars_numerical, aes(x = number_of_photos)) + geom_histogram(fill = "#00AFBB", colour="white", bins=40) + labs(title = 'Histogram of number_of_photos')

grid.arrange(plot7, plot8, plot9, plot10, ncol=2, nrow=2)
```

The Q-Q plots and histograms also show evidence of non-normality. The `odometer_value`, `engine_capacity` and `number_of_photos` are right-skewed, while `year_produced` is left-skewed.

Now we apply the *Kolmogorov-Smirnov normality test* into our data. The null hypothesis of this test is 'sample distribution is normal'.

```{r, warning=FALSE}
ks.test(df$price, 'pnorm', mean=mean(df$price), sd=sd(df$price))
ks.test(df$odometer_value, 'pnorm', mean=mean(df$odometer_value), sd=sd(df$odometer_value))
ks.test(df$year_produced, 'pnorm', mean=mean(df$year_produced), sd=sd(df$year_produced))
ks.test(df$engine_capacity, 'pnorm', mean=mean(df$engine_capacity), sd=sd(df$engine_capacity))
ks.test(df$number_of_photos, 'pnorm', mean=mean(df$number_of_photos), sd=sd(df$number_of_photos))
```

The p-value of all the numeric variables are < 2e-16 which is less than 0.05, therefore we conclude that the distributions of all our numeric variables are significantly different from normal distribution. They have the same results with Q-Q plots and histograms.

Our sample size for this data is 38521. Based on the central limit theorem, we generate the rest analysis using the original data.



## SMART Qeustions
The following are our SMART questions

**Specific**: Is it possible to build a model to find a relationship between car prices by looking at different factors that include numerical, categorical values and further use the model to predict car prices?

**Measurable**: Is it possible to measure metrics such as r-square, MAE, MSE and RMSE with your categories?

**Achievable**: Based on the preliminary analysis that the team concluded is it possible to find a pattern between target variable(car price) and the independent variable?

**Relevant**: Can the research help the sellers and buyers in the used car market to make an informed decision about the price of the vehicle?

**Time Oriented**: Will The initial analysis be completed by November, 2nd with the presentation?






## Correlation Plot

```{r Fig 1}
corrplot(cor(cars_numerical), method = 'number')
```
Figure 1 shows the correlation between the numerical features.

The team used a correlation plot for checking the correlation between continuous variables. Year of production was highly correlated with price with correlation coefficient(cc)=0.7. Odometer value had a negative correlation with year produced (cc=-0.49) and price (cc=-0.42). Engine capacity also had a positive correlation with price (cc=0.30). 


```{r Fig. 2}
library(ggplot2)

df %>% group_by(year_produced) %>% summarize(mean_price_per_year = mean(price, na.rm=TRUE)) %>% ggplot(aes(x=year_produced,y=mean_price_per_year)) +  geom_col(fill = "#00AFBB") + labs(title='Avg Price of Car per Year', x="year produced", y = "mean price per year") + theme(plot.title = element_text(hjust = 0.5))
```

Figure 2 shows the average price of the car for each year produced between 1940 and 2020. The team observed that there is a steady decrease in the price as the car gets older. However around 1990, it could be observed that the prices spike as cars before 1990 fall under the classic or vintage category.

The bar plot of the average price of the car in different years showed that the vintage cars produced around the year 1965 are pricier than the newer cars. And the price increased steadily after around 1985.


```{r Fig. 3}
df %>% group_by(engine_capacity) %>% summarize(mean_price_per_capicity = mean(price, na.rm=TRUE)) %>% ggplot(aes(x=engine_capacity,y=mean_price_per_capicity)) +  geom_point(color = "#00AFBB") + labs(title='Avg Price of Car for engine capacity', x='Engine Capacity', y='Mean Price') + theme(plot.title = element_text(hjust = 0.5))
```
Figure 3 shows the average price of the car for each engine capacity. 
The team observed a positive linear trend between the mean price per engine capacity and the capacity 



```{r}
df %>% group_by(engine_capacity) %>% summarize(mean_price_per_capacity = mean(price, na.rm=TRUE)) ->df4
cor(df4)
#corrplot(cor(cars_numerical), method = 'number')
```

The observed correlation coefficient equals 0.6. However, in Figure 1 it was observed that the correlation coefficient between price and engine capacity was 0.3. This trend could be explained by the outliers which are found in higher engine capacity.



```{r Fig 4}
df %>% group_by(engine_capacity) %>% summarize(mean_price_per_capacity = mean(price, na.rm=TRUE)) ->df4
cor(df4)
#corrplot(cor(cars_numerical), method = 'number')
```

```{r}
df %>% ggplot(aes(x=reorder(body_type,-engine_capacity),y=engine_capacity, fill=body_type))+geom_boxplot() + labs(x='Body Type', y='Engine Capicity')  + ggtitle('Body Type vs Engine Capicity ') + theme(plot.title = element_text(hjust = 0.5))
```

Figure 4 shows the mean engine capacity for different body type using a box-plot. From the initial analysis the team observed for each of the groups there is a difference in median.

## Mean Difference Between Variables
### Warranty vs Price
```{r}
df %>% ggplot(aes(has_warranty, price, fill=has_warranty)) + geom_boxplot() + ggtitle('Has_Warranty vs Prices ') + theme(plot.title = element_text(hjust = 0.5))
```

The graph indicates a substantial difference of mean between the two subsets
Let's run a t-test to check the statistical significance 
```{r T test, echo=TRUE}
summary(df$has_warranty)
has = subset(df, has_warranty == "True")
hasnot = subset(df, has_warranty == "False")
t.test(x = has$price, y = hasnot$price, conf.level = 0.99)
```
The t-test on the has_warranty feature suggest strong statistical significance and thus we can reject null hypothesis, suggesting that there is a difference between the true means of the two subsets.


###  Engine Types vs Price
```{r}
df %>% ggplot(aes(engine_type, price,fill=engine_type)) + geom_boxplot()+ ggtitle('Engine_type vs Prices ') + theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=TRUE}
summary(df$engine_type)
diesel = subset(df, subset = df$engine_type == "diesel")
gas = subset(df, subset = df$engine_type == "gasoline")
t.test(x = diesel$price, y = gas$price, conf.level = 0.99)
```

The t-test suggest there is a difference between the true means of the two subsets(Diesel,Gas).



## Independence and Homogeneity
```{r}
contgcTbl1 = table(df$manufacturer_name, df$has_warranty)

(Xsq1 = chisq.test(contgcTbl1))


contgcTbl2 = table(df$manufacturer_name, df$body_type)

(Xsq2 = chisq.test(contgcTbl2))

contgcTbl3 = table(df$manufacturer_name, df$color)

(Xsq3 = chisq.test(contgcTbl3))

(contgcTbl4 = table(df$color, df$transmission))

(Xsq4 = chisq.test(contgcTbl4))

(contgcTbl5 = table(df$manufacturer_name, df$is_exchangeable))

(Xsq5 = chisq.test(contgcTbl5))


```
For test of independence homogeneity, we tested manufacturer names against whether has warranty, the body type of cars, car colors and whether the car is exchangable, which are for table 1, 2 , 3 and 5. We also tested car colors against transmissions, which is table 4.

For table 1, 2 and 3, because many of the expected values are so small, the approximation of p may not be right.


## Influential Feactures of Prices
### Colors by Mean Price
```{r}
df %>% group_by(color) %>% summarise(price_colorMean=mean(price)) %>% ggplot(aes(x=reorder(color,-price_colorMean),y=price_colorMean)) + geom_col(fill = "#00AFBB") + labs(x='Color',y='Price mean') + ggtitle('Color vs Prices ') + theme(plot.title = element_text(hjust = 0.5))
```

### Body Types by Mean Price
```{r}
df %>% group_by(body_type) %>% summarise(body_price_mean = mean(price))%>% ggplot(aes(x = reorder(body_type, -body_price_mean),body_price_mean))+geom_col(fill = "#00AFBB") + labs(x='Body Type', y='Mean of price') + ggtitle('Body Type vs Price ') + theme(plot.title = element_text(hjust = 0.5))
```

### Top 10 Manufacturers by Mean Price
```{r}
df2 = df %>% group_by(manufacturer_name) %>% summarise(manuf_price_mean = mean(price)) %>% arrange(desc(manuf_price_mean)) 
df2 %>% slice(1:10) %>%  ggplot(aes(x = reorder(manufacturer_name, -manuf_price_mean),manuf_price_mean))+geom_col(fill = "#00AFBB") + labs(x='Manufacturer', y='Mean of price')  + ggtitle('Manufacturer vs Price ') + theme(plot.title = element_text(hjust = 0.5))
```

From the initial analysis, for the above 3 plots there seems to be difference among the true means of their respective groups. Lets run an anova test on each of the case to check the hypothesis


### One Way ANOVA
```{r}
df_aov_1 = aov(price ~  color , df)
summary(df_aov_1)

df_aov_2 = aov(price ~  manufacturer_name , df)
summary(df_aov_2)

df_aov_2 = aov(price ~  body_type , df)
summary(df_aov_2)

```

For all the three cases there seems to be statistically significant difference the the true means
The high statistical significance suggest that there is a difference between the true mean of the prices for differently colored vehicles, for different manufacturer and for different body types



### Two Way ANOVA
```{r}
df_aov = aov(price ~  manufacturer_name * body_type , df)
summary(df_aov)

#Check for homoscedasticity 
#par(mfrow=c(2,2))
#plot(df_aov)
#par(mfrow=c(1,1))
```

The high statistical significance show that the interaction term can explain the variation in the price variable.



```{r}
df_aov = aov(price ~ manufacturer_name * color, df)
summary(df_aov)

#TukeyHSD(df_aov)

```
Because the p < 0.01, we reject that all colors sell the same price, all manufacturer sell the same price and manufacturers and colors combined sell the same price.



```{r pressure, echo=FALSE}
#boxplot(df['manufacturer_name'], df['price'])
#car_lm = lm(price ~ ., data = df2)


#library(s20x)
#pairs20x(df2)
```
```{r}
#library(ezids)
#loadPkg("corrplot")
#corrplot(df2)
```


## Conclusion and Discussions
Overall, our work involved removing the null values for data preprocessing, data exploratory, normality check, finding the correlation between continuous variables, and finding the mean price difference between multiple categorical variables. The technologies we used included a table summary, normality tests, t-test, ANOVA, and Chi-square test. We used a variety of plots such as bar plot, scatter plot, box plot, Q-Q plot, and histogram to support different tests.

For more details, we deleted ten null values in the data preprocessing part. Then we generated a table to show the basic statistical measurements of numeric data. The price of this data offers mean=6640 and standard deviation=6430. The other two measurements we may consider are skewness and kurtosis. These two statistical values indicated that the data were highly skewed.

Based on these results, we checked the normality of continuous data by using Q-Q plot, histogram, and Kolmogorov-Smirnov normality test. The normality tests showed significant evidence to reject the null hypothesis. Thus, the price was not a normal distribution. The other continuous variables showed the same results. Therefore, for our future work, if we need to use price as the dependent variable to create a regression, we will transform the data to a normal distribution.

We used a correlation plot for checking the correlation between continuous variables. Year of production was highly correlated with price with correlation coefficient(cc)=0.7. Odometer value had a negative correlation with year produced (cc=-0.49) and price (cc=-0.42). Engine capacity also had a positive correlation with price (cc=0.30).

We then generated other exploratory data analysis for the feature that we were more concerned about – price. 

The bar plot of the average price of the car in different years showed that the vintage cars produced around the year 1965 are pricier than the newer cars. And the price increased steadily after around 1985. The box plots and t-tests suggested the solid statistical significance of the difference between the mean price of vehicles with a warranty and without warranty and diesel and gasoline engine types. In our analysis, one-way and two-way ANOVA were used to check the difference between more than three levels of categorical data and price. The results suggested that color, manufacturer name, and body type had mean price differences.


According to the above analysis, the features that influence the prices of cars in the used car market in Belarus are year of production, body type, manufacture name, engine capacity, odometer value, engine type color, and transmission. 

After conducting the EDA and hypothesis tests on the data, the team has concluded that the initial SMART research question were successful answered.

Our future work for this topic is building up a model to predict the price based on the analysis we explored to provide more effective decision-making services for future vehicle buyers and sellers.


